---
title: "HW5-Group7-Cohort2"
author: "Cohort 2, Group 7 - Hyeuk Jung, Jiaqi Li, Xichen Luo, Huanyu Liu"
date: "March 3, 2019"
output: pdf_document
---

# Principal Component Analysis
```{r}
library(data.table)
library(dplyr)
library(sandwich)
library(ggplot2)

industry_48 <- read.csv("48_Industry_Portfolios_vw.csv", header = T) %>% as.data.table; gc()
famafrench <- read.csv("F-F_Research_Data_Factors.csv", header = T) %>% as.data.table; gc()

# Data Cleaning and Manipulation
# 1. Subset the data: year 1960~2015
industry_48$year <- as.numeric(substr(industry_48$X, 1, 4)) 
industry <- industry_48[ (industry_48$year >= 1960) & (industry_48$year < 2016), ] 
# 2. Remove columns if it holds -99.99 or missing values
# changing -99.99 values to NA
industry[industry == -99.99] <- NA 
industry[industry == -999] <- NA
# removing columns with NAs
industry_filtered <- industry %>% select_if( ~ !any(is.na(.)) ) 
# 3. Merge industry data and Rf from Fama-French data
data <- merge(industry_filtered, famafrench, by = "X")
```

## Problem 1

```{r}
# 1. Calculate the excess return of each industry
ex.ret <- (data[, 2:44] - data$RF) #/100 # change into percentile dim(ex.ret)
# 2. Get the var-cov matrix
varcov <- cov(ex.ret) 
# 3. Get the eigenvalues
eigenvalues <- eigen(varcov) #$values #%>% as.data.table; gc()
# 4. Plot the fraction of variance explained by each eigenvalue
var <- sum(eigenvalues$values)
var_fraction <- eigenvalues$values / var
var_fraction_plot <- barplot(var_fraction, 
                             main = "Fraction of Variance Explained by Eigenvalues", 
                             ylab = "Fraction of Variance", xlab = "Industry", 
                             ylim = c(0, 0.7))
```

## Problem 2

#### (a)
```{r}
# 2.(a). Explanation power of these three components
sum(var_fraction[1:3]) 
```

#### Around 69.88% of the total variance is explained by these 3 factors.

#### (b)
```{r}
prcomp_result <- prcomp(ex.ret)
summ_prcomp_result <- summary(prcomp_result) #summ_prcomp_result$rotation
# Using prcomp results (did not find the PC1 ~ PC3 values from princomp result)
largest_pca <- as.data.frame(prcomp_result$rotation[, 1:3])
# monthly returns for each industry*loadings
factor_return <- as.matrix(ex.ret) %*% as.matrix(largest_pca) 
facor_return_mean <- apply(factor_return, 2, mean)
```

#### Mean sample return to these 3 factor portfolios:
```{r}
facor_return_mean
```

#### Standard deviation of these 3 factors portfolios
```{r}
prcomp_result$sdev[1:3] 
```

#### Correlation of these 3 factors portfolios
```{r}
cor(factor_return)
```

#### Factor loadings
```{r}
ggplot(data = largest_pca) + 
  geom_line(aes(x = 1:length(largest_pca$PC1), y = PC1), color = "red") +
  geom_line(aes(x = 1:length(largest_pca$PC2), y = PC2), color = "blue") +
  geom_line(aes(x = 1:length(largest_pca$PC3), y = PC3), color = "black") +
  labs(title = "PC1~3", x = "Industry", y = "PC")
```

#### (c)
```{r, fig.width=6.5, fig.height=4.3}
Re_real_mean <- apply(ex.ret, 2, mean) # Realized avg

y <- as.matrix(ex.ret)
X <- solve(prcomp_result$rotation) %*% t(y)
X1 <- as.matrix(X[1, ])
X2 <- as.matrix(X[2, ])
X3 <- as.matrix(X[3, ])
T <- length(y[, 1]) # 672 dim(X1%*%t(as.matrix(largest_pca[, 1])))
F3 <- X1%*%t(as.matrix(largest_pca[, 1])) + X2%*%t(as.matrix(largest_pca[, 2])) + 
  X3%*%t(as.matrix(largest_pca[, 3]))
F3_mean <- apply(F3, 2, mean)

betas <- t(largest_pca) # betas for each industry #dim(largest_pca) -> 43by3; dim(betas) -> 3by43
Re_predict <- factor_return %*% betas # dim(factor_return) -> 672by3 x dim(betas) -> 3by43
Re_predict_mean <- apply(Re_predict, 2, mean) #dim(betas) dim(factor_return) -> 672by3

plot(x = F3_mean, y = Re_real_mean, pch = 19, col = "red", ylim = c(0, 1), 
     xlim = c(0, 1), xlab = NA, ylab = NA)
abline(0, 1)
par(new=TRUE)
plot(x = Re_predict_mean, y = Re_real_mean, pch = 20, col = "blue", 
     ylim = c(0, 1), xlim = c(0, 1), xlab="Red: F3_mean; Blue: Re_predict_mean")
abline(0, 1)

final <- cbind(Re_real_mean, Re_predict_mean, F3_mean)
ggplot(data = as.data.frame(final)) + 
  geom_line(aes(x = 1:length(final[, 1]), y = Re_real_mean), color = "red") +
  geom_line(aes(x = 1:length(final[, 2]), y = Re_predict_mean), color = "blue") +
  geom_line(aes(x = 1:length(final[, 3]), y = F3_mean), color = "black") +
  labs(title = "Mean Returns", x = "Industry", y = "Returns")
```


#### (d)
```{r}
rsquard_1 <- 1 - var(Re_real_mean - F3_mean) / var(Re_real_mean)
rsquard_2 <- 1 - var(Re_real_mean - Re_predict_mean) / var(Re_real_mean)

rsquard_1
rsquard_2
```

## Problem 3

```{r}
industry_25 <- read.csv("25_Portfolios_5x5_vw.csv", header = T) %>% as.data.table; gc()
# Data Cleaning and Manipulation
# 1. Subset the data: year 1960~2015
# get year info of each row
industry_25$year <- as.numeric(substr(industry_25$X, 1, 4)) 
# year condition
industry_2 <- industry_25[ (industry_25$year >= 1960) & (industry_25$year < 2016), ] 
# 2. Remove columns if it holds -99.99 or missing values
industry_2[industry_2 == -99.99] <- NA # changing -99.99 values to NA
industry_2[industry_2 == -999] <- NA
# removing columns with NAs
industry_2_filtered <- industry_2 %>% select_if( ~ !any(is.na(.)) ) 
# 3. Merge industry data and Rf from Fama-French data
data_2 <- merge(industry_2_filtered, famafrench, by = "X")
```

#### (a)
```{r, fig.width=6.5, fig.height=4.2}
# 1. Calculate the excess return of each industry
ex.ret_2 <- (data_2[, 2:26] - data$RF) #/100 # change into percentile
dim(ex.ret_2)
# 2. Get the var-cov matrix
varcov_2 <- var(ex.ret_2) 
# 3. Get the eigenvalues
eigenvalues_2 <- eigen(varcov_2) #$values #%>% as.data.table; gc()
# 4. Plot the fraction of variance explained by each eigenvalue
var_2 <- sum(eigenvalues_2$values)
var_fraction_2 <- eigenvalues_2$values / var_2
var_fraction_plot_2 <- barplot(var_fraction_2, main = "Fraction of Variance Explained by Eigenvalues", 
                             ylab = "Fraction of Variance", xlab = "Industry",
                             ylim = c(0, 1))
```

#### (b)
```{r}
summary(prcomp(ex.ret_2)) 
```
#### According to the plot in (a), PC1 covers the largest fraction of variance, following are PC2, PC3, and PC4. In that case, 4 factors needed to explain average returns to the 25 F-F portfolios